---
title: "Draft Stress test comparison"
output:
  html_document: default
  pdf_document: default
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=8, fig.height=6) 
```

Load functions.

```{r}
devtools::load_all()
```

Load libraries

```{r}
library(patchwork)
```

Read in the first data set. Put the outputs of a stress test run into 
'data/data1' folder of this project to run the code successfully.

```{r}
st_output_path1 <- here::here("data/data1")

expected_loss_file <- here::here(st_output_path, "loans_company_expected_loss_standard.csv")
data_expected_loss1 <- read.csv(expected_loss_file)
# Expected loss data preparation. Initial filtering and assigning variables to plot. The values are turned into millions.
data_el1 <- prepare_for_el_plots(data_expected_loss)

pd_file <- here::here(st_output_path, "loans_portfolio_pd_changes_overall_standard.csv")
data_pd <- read.csv(pd_file)
data_pd1 <- prepare_for_pd_plot(data_pd)

value_changes_file <- here::here(st_output_path, "loans_portfolio_value_changes_standard.csv")
data_val_change1 <- read.csv(value_changes_file)
```

Read in the second data set. Put the outputs of a stress test run into 
'data/data1' folder of this project to run the code successfully.

```{r}
st_output_path1 <- here::here("data/data2")

expected_loss_file <- here::here(st_output_path, "loans_company_expected_loss_standard.csv")
data_expected_loss2 <- read.csv(expected_loss_file)
# Expected loss data preparation. Initial filtering and assigning variables to plot. The values are turned into millions.
data_el2 <- prepare_for_el_plots(data_expected_loss)

pd_file <- here::here(st_output_path, "loans_portfolio_pd_changes_overall_standard.csv")
data_pd <- read.csv(pd_file)
data_pd2 <- prepare_for_pd_plot(data_pd)

value_changes_file <- here::here(st_output_path, "loans_portfolio_value_changes_standard.csv")
data_val_change2 <- read.csv(value_changes_file)
```

## PD changes comparison

```{r}
p1 <- qplot_pd_sector(data_pd1)
p2 <- qplot_pd_sector(data_pd2)

p1 / p2
```
## EL changes comparison 

Percentage

```{r}
p1 <- qplot_el_change_sector_perc(data_el1)
p2 <- qplot_el_change_sector_perc(data_el2)

p1 / p2
```

Absolute values

```{r}
p1 <- qplot_el_change_sector_abs(data_el1)
p2 <- qplot_el_change_sector_abs(data_el2)

p1 / p2

```
## Equity value change comparison

Percentage sector

```{r}
p1 <- qplot_val_change_sector_perc(data_val_change1)
p2 <- qplot_val_change_sector_perc(data_val_change2)

p1 / p2
```

Absolute technology

```{r fig, fig.width = 8, fig.height = 10}
p1 <- qplot_val_change_technology_abs(data_val_change1)
p2 <- qplot_val_change_technology_abs(data_val_change2)

p1 / p2
```
