---
title: "Draft ST plots"
output:
  html_document: default
  pdf_document: default
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=8, fig.height=6) 
```

Load functions.

```{r}
devtools::load_all()
```

Read in the data. Put the outputs of a stress test run into 'data' folder of this project to run the code successfully.

```{r}
st_output_path <- here::here("data", "initial repo")

expected_loss_file <- here::here(st_output_path, "loans_company_expected_loss_standard.csv")
data_expected_loss <- read.csv(expected_loss_file)
# Expected loss data preparation. Initial filtering and assigning variables to plot. The values are turned into millions.
data_plot <- prepare_for_exposure_plot(data_expected_loss)
data_summary <- prepare_for_summary_plot(data_expected_loss)
data_el <- prepare_for_el_plots(data_expected_loss)

pd_file <- here::here(st_output_path, "loans_portfolio_pd_changes_overall_standard.csv")
data_pd <- read.csv(pd_file)
data_pd <- prepare_for_pd_plot(data_pd)

pd_annual_file <- here::here(st_output_path, "loans_portfolio_pd_changes_annual_standard.csv")
data_pd_annual <- read.csv(pd_annual_file)
data_pd_annual <- prepare_for_pd_annual_plot(data_pd_annual) 

value_changes_file <- here::here(st_output_path, "loans_portfolio_value_changes_standard.csv")
data_val_change <- read.csv(value_changes_file)
```

# Exposure plots

## Summary exposures per sector

Pie chart of exposures summary

```{r}
qplot_pie_summaries(data_summary)
```

## Risk-driving companies (based on exposure)

10 worst performers - coloured by sector

```{r fig2, fig.width = 8, fig.height=8}
qplot_worst_performers(data_plot)
```

## Distributions

### Histograms and densities per sector

```{r}
qplot_histogram_density(data_plot) 
```


### Boxplots and violin plots

Boxplots

```{r}
qplot_boxplot(data_plot)
```

Points distribution with violin

```{r}
qplot_violin(data_plot)
```

## PD changes

### Overall

Faceted - joint y-axis (easier to compare between sectors)

```{r}
qplot_pd_sector(data_pd)
qplot_pd_sector(data_pd, common_y_axis = TRUE)
```

### Annual

Faceted - joint y-axis (easier to compare between sectors)

```{r}
qplot_pd_sector(data_pd_annual, annual_pds = TRUE)
qplot_pd_sector(data_pd_annual, annual_pds = TRUE, common_y_axis = TRUE)
```


## EL changes

```{r fig3, fig.width = 8, fig.height = 8} 
qplot_el_sector(data_el)
```
```{r}
qplot_el_change_sector(data_el)
```

## Equity Value changes

Total

```{r}
qplot_val_change(data_val_change)
```

Per sector

```{r} 
qplot_val_change_sector(data_val_change)
```

Per technology

```{r fig4, fig.width = 8, fig.height = 12}
qplot_val_change_technology(data_val_change)
```
