---
title: "Draft ST plots"
output:
  html_document: default
  pdf_document: default
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=8, fig.height=6) 
```

Load functions.

```{r}
devtools::load_all()
```

Read in the data. Put the outputs of a stress test run into 'data' folder of this project to run the code successfully.

```{r}
st_output_path <- here::here("data")

expected_loss_file <- here::here(st_output_path, "loans_company_expected_loss_standard.csv")
data_expected_loss <- read.csv(expected_loss_file)
# Expected loss data preparation. Initial filtering and assigning variables to plot. The values are turned into millions.
data_plot <- prepare_for_exposure_plot(data_expected_loss)
data_summary <- prepare_for_summary_plot(data_expected_loss)
data_el <- prepare_for_el_plots(data_expected_loss)

pd_file <- here::here(st_output_path, "loans_portfolio_pd_changes_overall_standard.csv")
data_pd <- read.csv(pd_file)
data_pd <- prepare_for_pd_plot(data_pd)

pd_annual_file <- here::here(st_output_path, "loans_portfolio_pd_changes_annual_standard.csv")
data_pd_annual <- read.csv(pd_annual_file)
data_pd_annual <- prepare_for_pd_annual_plot(data_pd_annual) 

value_changes_file <- here::here(st_output_path, "loans_portfolio_value_changes_standard.csv")
data_val_change <- read.csv(value_changes_file)
```

# Exposure plots

## Summary exposures per sector

Pie chart of exposures summary

```{r}
qplot_pie_summaries(data_summary)
```

## Risk-driving companies (based on exposure)

10 worst performers - coloured by sector

```{r fig2, fig.width = 8, fig.height=8}
qplot_worst_performers(data_plot)
```

## Distributions

### Histograms and densities per sector

```{r}
qplot_histogram_density(data_plot) 
```


### Boxplots and violin plots

Boxplots

```{r}
qplot_boxplot(data_plot)
```

Points distribution with violin

```{r}
qplot_violin(data_plot)
```

## PD changes

### Overall

Faceted - joint y-axis (easier to compare between sectors)

```{r}
qplot_pd_sector(data_pd)
qplot_pd_sector(data_pd, common_y_axis = TRUE)
```

### Annual

Faceted - joint y-axis (easier to compare between sectors)

```{r}
qplot_pd_sector(data_pd_annual, annual_pds = TRUE)
qplot_pd_sector(data_pd_annual, annual_pds = TRUE, common_y_axis = TRUE)
```


## EL changes

```{r} 
data_losses <- data_el %>%
  filter(name %in% c("expected_loss_baseline", "expected_loss_shock")) %>%
  mutate(
    name = str_replace(name, "expected_loss_", "")
  )

ggplot(data_losses, aes(x = name, y = value, fill = value)) +
  geom_bar(stat = "identity", color = "grey") +
  scale_x_discrete(position = "top", labels = r2dii.plot::to_title) +
  scale_y_continuous(expand = expansion(mult = c(.1, 0)), labels = comma) +
  scale_fill_gradient2(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "green") %>% pull(.data$hex),
      midpoint = 0,
      labels = comma,
      name = "Expected loss\n(currency)"
    ) +
  facet_wrap(~ ald_sector, scales = "free") +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    strip.placement = "outside"
  ) +
  labs(
    title = "Expected loss per sector",
    subtitle = "For baseline and shock scenarios",
    y = "Expected Loss (currency)"
  )
```
```{r}
ggplot(data_losses, aes(x = name, y = el_as_perc_exposure, fill = el_as_perc_exposure)) +
  geom_bar(stat = "identity", color = "grey") +
  scale_x_discrete(position = "top", labels = r2dii.plot::to_title) +
  scale_y_continuous(expand = expansion(mult = c(.1, 0)), labels = percent) +
  scale_fill_gradient2(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "green") %>% pull(.data$hex),
      midpoint = 0,
      labels = percent,
      name = "Expected loss\n(% exposure)"
    ) +
  facet_wrap(~ ald_sector, scales = "free") +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    strip.placement = "outside"
  ) +
  labs(
    title = "Expected Loss as percentage of exposure per sector",
    subtitle = "For baseline and shock scenarios",
    y = "Expected Loss (% exposure)"
  )
```
```{r}
data_change <- data_el %>%
  filter(name %in% c("value_change")) 

p_abs <- ggplot(data_change, aes(x = name, y = value, fill = value)) +
  geom_bar(stat = "identity", color = "grey") +
  scale_x_discrete(position = "top", labels = r2dii.plot::to_title) +
  scale_y_continuous(expand = expansion(mult = c(.1, 0)), labels = comma) +
  scale_fill_gradient2(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "green") %>% pull(.data$hex),
      midpoint = 0,
      labels = comma,
      name = "Expected loss\n(currency)"
    ) +
  facet_wrap(~ ald_sector) +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    strip.placement = "outside"
  ) +
  labs(
    title = "Expected loss shock per sector",
    y = "Expected Loss\n(currency)"
  )

p_perc <- ggplot(data_change, aes(x = name, y = el_as_perc_exposure, fill = el_as_perc_exposure)) +
  geom_bar(stat = "identity", color = "grey") +
  scale_x_discrete(position = "top", labels = r2dii.plot::to_title) +
  scale_y_continuous(expand = expansion(mult = c(.1, 0)), labels = percent) +
  scale_fill_gradient2(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "green") %>% pull(.data$hex),
      midpoint = 0,
      labels = percent,
      name = "Expected loss\n(% exposure)"
    ) +
  facet_wrap(~ ald_sector) +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    strip.placement = "outside"
  ) +
  labs(
    title = "Expected loss shock as percentage of exposure per sector",
    y = "Expected Loss\n(% exposure)"
  )

p_abs / p_perc
```

## Equity Value changes

Total

```{r}
qplot_val_change(data_val_change)
```

Per sector

```{r} 
qplot_val_change_sector(data_val_change)
```

Per technology

```{r fig3, fig.width = 8, fig.height = 12}
qplot_val_change_technology(data_val_change)
```
