---
title: "Draft ST plots"
output:
  html_document: default
  pdf_document: default
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=8, fig.height=6) 
```

# Data preparation

Call to libraries

```{r message = FALSE, warning = FALSE}
library(here)
library(dplyr)
library(tidyr)
library(tidytext)
library(ggplot2)
library(r2dii.plot)
library(r2dii.colours)
library(patchwork)
library(forcats)
library(lubridate)
library(scales)
library(stringr)
```

Read in the data. Put the outputs of a stress test run into 'data' folder of this project to run the code successfully.

```{r}
st_output_path <- here("data")

expected_loss_file <- here(st_output_path, "loans_company_expected_loss_standard.csv")
data_expected_loss <- read.csv(expected_loss_file)

pd_file <- here(st_output_path, "loans_portfolio_pd_changes_overall_standard.csv")
data_pd <- read.csv(pd_file)

pd_annual_file <- here(st_output_path, "loans_portfolio_pd_changes_annual_standard.csv")
data_pd_annual <- read.csv(pd_annual_file)

value_changes_file <- here(st_output_path, "loans_portfolio_value_changes_standard.csv")
data_val_change <- read.csv(value_changes_file)
```

Expected loss data preparation. Initial filtering and assigning variables to plot. The values are turned into millions.

```{r}
data_plot <- data_expected_loss %>%
  filter(
    scenario_name == "Carbon balance 2030",
    scenario_geography == "Global",
    investor_name == "Meta Investor",
    portfolio_name == "Meta Investor"
  ) %>%
  mutate(
    value_to_plot = .data$exposure_at_default,
    group_variable = .data$ald_sector
    ) %>%
  mutate(
    value_to_plot = .data$value_to_plot / 10^6
  ) %>%
  group_by(.data$group_variable) %>%
  mutate(
    group_mean = mean(.data$value_to_plot, na.rm = TRUE),
    group_sum = sum(.data$value_to_plot, na.rm = TRUE),
    group_max = max(.data$value_to_plot, na.rm = TRUE)
  )

data_summary <- data_plot %>%
  group_by(.data$group_variable) %>%
  summarise(
    group_sum = sum(.data$value_to_plot, na.rm = TRUE),
    group_n = n()
  ) %>%
  pivot_longer(cols = c("group_sum", "group_n")) %>%
  mutate(
    label = case_when(
      .data$name == "group_sum" ~ "Total exposure per sector (Millions)",
      .data$name == "group_n" ~ "Number of companies per sector"
    )
  ) %>%
  group_by(name) %>%
  mutate(
    perc_of_total = .data$value / sum(.data$value)
  )
```

PD data preparation.

```{r}
data_pd <- data_pd %>%
  filter(
    scenario_geography == "Global",
    investor_name == "Meta Investor",
    portfolio_name == "Meta Investor"
  ) %>%
  select(ald_sector, term, shock_year_arg, pd_change_sector_shock) %>%
  distinct()

data_pd_annual <- data_pd_annual %>%
  filter(
    scenario_geography == "Global",
    investor_name == "Meta Investor",
    portfolio_name == "Meta Investor"
  ) %>% 
  select(ald_sector, year, shock_year_arg, pd_change_sector_shock) %>%
  distinct() %>%
  mutate(
    year_date = ymd(year, truncated = 2L)
  )
```

EL data preparation
```{r}
data_exp_per_sector <- data_summary %>%
  filter(
    name == "group_sum"
  ) %>%
  ungroup() %>%
  select(-name) %>%
  mutate(
    ald_sector = group_variable, 
    exposure = value*10^6
    ) %>%
  select(ald_sector, exposure)

data_el <- data_expected_loss %>%
  filter(
    scenario_geography == "Global",
    investor_name == "Meta Investor",
    portfolio_name == "Meta Investor"
  ) %>%
  group_by(ald_sector) %>%
  summarise(
    expected_loss_baseline = -sum(expected_loss_baseline, na.rm = TRUE),
    expected_loss_shock = -sum(expected_loss_shock, na.rm = TRUE),
    value_change = expected_loss_shock - expected_loss_baseline
  ) %>%
  pivot_longer(
    cols = c(expected_loss_baseline, expected_loss_shock, value_change)
  ) %>%
  inner_join(data_exp_per_sector, by = "ald_sector") %>%
  mutate(
    el_as_perc_exposure = value / exposure
  )
```

# Exposure plots

## Summary exposures per sector

Pie chart of exposures summary

```{r}
ggplot(data_summary, aes(x = "", y = .data$perc_of_total, fill = .data$group_variable)) +
  geom_bar(stat = "identity", width = 1) +
  scale_fill_2dii("pacta", colour_groups = data_summary$group_variable) +
  coord_polar("y", start = 0) +
  theme_2dii() +
  theme(
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank()
  ) +
  facet_wrap(~label) +
  
  geom_text(aes(x = 1.6, label = round(.data$value)), position = position_stack(vjust = .5)) +
  
  labs(
    title = "Portfolio composition"
  )
```

## Risk-driving companies (based on exposure)

10 worst performers - coloured by sector

```{r fig2, fig.width = 8, fig.height=8}
data_worst_performers <- data_plot %>%
  group_by(.data$group_variable) %>%
  slice_max(order_by = .data$value_to_plot, n = 10) %>%
  ungroup() %>%
  mutate(
    company_name = reorder_within(.data$company_name, .data$value_to_plot, .data$group_variable)
    )

ggplot(data_worst_performers, aes(x = .data$company_name, y = .data$value_to_plot, fill = .data$group_variable)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(.data$value_to_plot, 1)), hjust = -0.2) +
  scale_x_reordered() +
  scale_y_continuous(expand = expansion(mult = c(0, .1))) +
  scale_fill_2dii("pacta", colour_groups = data_worst_performers$group_variable) +
  coord_flip() +
  theme_2dii() +
  theme(
    axis.title.y = element_blank(),
    legend.position = "none"
  ) +
  facet_wrap(~fct_reorder(.data$group_variable, .data$value_to_plot, .fun = max, .desc = TRUE), scales = "free_y", ncol = 1) +
  labs(
    title = "Companies with the highest exposure within sector",
    y = "Exposure (Millions)"
  )
```

## Distributions

### Histograms and densities per sector

```{r}
phist <- ggplot(data_plot, aes(x = .data$value_to_plot, fill = .data$group_mean)) +
  geom_histogram(bins = 15) +
  scale_y_continuous(expand = expansion(mult = c(0, .1))) +
  scale_fill_gradient(
    low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "grey") %>% pull(.data$hex),
    high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
    name = "Average exposure\n per sector (millions)"
  ) +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    axis.title.x = element_blank()
  ) +
  labs(
    title = "Distributions of exposures per sector",
    subtitle = "Demonstrated using histograms and density plots.\n The colours represent the average size of the exposure within a data set.",
    x = "Exposure (millions)",
    y = "Number of companies"
  ) +
  facet_wrap(~.data$group_variable, scales = "free")

pdens <- ggplot(data_plot, aes(x = .data$value_to_plot, fill = .data$group_mean)) +
  geom_density(aes(y = ..scaled..)) +
  scale_y_continuous(expand = expansion(mult = c(0, .1))) +
  scale_fill_gradient(
    low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "grey") %>% pull(.data$hex),
    high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
    name = "Average exposure\n per sector (millions)"
  ) +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    strip.text = element_blank()
  ) +
  labs(
    x = "Exposure (millions)",
    y = "Density"
  ) +
  facet_wrap(~.data$group_variable, scales = "free")

(phist / pdens + plot_layout(guides = "collect")) 
```


### Boxplots and violin plots

Outliers removal

```{r}
data_plot <- data_plot %>%
        group_by(.data$group_variable) %>%
        mutate(
          is_outlier = (abs(.data$value_to_plot - median(.data$value_to_plot)) > 2 * sd(.data$value_to_plot, na.rm = TRUE))
          ) 

data_plot %>% summarise(is_outlier = sum(is_outlier)) %>%
  glue::glue_data("{is_outlier} outliers are removed from {group_variable} sector data.")

data_plot <- data_plot %>%
  filter(!is_outlier) %>%
  select(-is_outlier)
```

Boxplots

```{r}
ggplot(data_plot, aes(x = .data$group_variable, y = .data$value_to_plot, fill = .data$group_mean)) +
  geom_boxplot() +
  scale_fill_gradient(
    low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "grey") %>% pull(.data$hex),
    high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
    name = "Average exposure\n per sector (millions)"
  ) +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    strip.text = element_blank()
  ) +
  labs(
    title = "Distributions of exposures per sector",
    subtitle = "Demonstrated using boxplots. The box shows the Interquartile range\n with black thick line showing the median. Black dots represent the outliers.\n The colours represent the average size of the exposure within a data set.",
    x = "Sector",
    y = "Exposure (millions)"
  ) +
  facet_wrap(~.data$group_variable, scales = "free") 
```

Points distribution with violin

```{r}
ggplot(data_plot, aes(x = .data$group_variable, y = .data$value_to_plot, colour = .data$value_to_plot)) +
  geom_violin() +
  geom_jitter(alpha = 0.5) + 
  scale_colour_gradient(
    low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "grey") %>% pull(.data$hex),
    high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
    name = "Exposure (millions)"
  ) +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    strip.text = element_blank()
  ) +
  labs(
    title = "Distributions of exposures per sector",
    subtitle = "Points represent a company. Position on the y-axis and colour represent\nthe size of the exposure. Curved lines approximate the shape of the distribution.",
    x = "Sector",
    y = "Exposure (millions)"
  ) +
  facet_wrap(~.data$group_variable, scales = "free")
```

## PD changes

### Overall

Faceted - joint y-axis (easier to compare between sectors)

```{r}
ggplot(data_pd, aes(x = term, y = pd_change_sector_shock, fill = pd_change_sector_shock)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(expand = expansion(mult = c(0, .1))) +
  scale_fill_gradient(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "grey") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      name = "PD change (% points)"
    ) +
  theme_2dii() +
  theme(
    legend.title = element_text()
  ) +
  facet_grid(shock_year_arg ~  ald_sector, scales = "free") +
  labs(
    title = "Overall PD change by sector and shock year",
    x = "Maturity (years)",
    y = "PD change (% points)"
    ) 
```

Faceted - separate y-axis (easier to see the PD change within sector)

```{r}
ggplot(data_pd, aes(x = term, y = pd_change_sector_shock, fill = pd_change_sector_shock)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(expand = expansion(mult = c(0, .1))) +
  scale_fill_gradient(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "grey") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      name = "PD change (% points)"
    ) +
  theme_2dii() +
  theme(
    legend.title = element_text()
  ) +
  facet_wrap(ald_sector ~ shock_year_arg, scales = "free") +
  labs(
    title = "Overall PD change by sector and shock year",
    x = "Maturity (years)",
    y = "PD change (% points)"
    ) 
```

### Annual

Faceted - joint y-axis (easier to compare between sectors)

```{r}
ggplot(data_pd_annual, aes(x = year_date, y = pd_change_sector_shock, fill = pd_change_sector_shock)) +
  geom_bar(stat = "identity") +
  scale_x_date() +
  scale_y_continuous(expand = expansion(mult = c(0, .1)), labels = comma) +
  scale_fill_gradient(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "grey") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      name = "PD change\n(% points)"
    ) +
  theme_2dii() +
  theme(
    legend.title = element_text()
  ) +
  facet_grid(shock_year_arg ~ ald_sector, scales = "free") +
  labs(
    title = "Annual PD changes per sector and shock year",
    x = "Year",
    y = "PD change (% points)"
    ) 
```

Faceted - separate y-axis (easier to see the PD change within sector)

```{r}
ggplot(data_pd_annual, aes(x = year_date, y = pd_change_sector_shock, fill = pd_change_sector_shock)) +
  geom_bar(stat = "identity") +
  scale_x_date() +
  scale_y_continuous(expand = expansion(mult = c(0, .1)), labels = comma) +
  scale_fill_gradient(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "grey") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      name = "PD change\n(% points)"
    ) +
  theme_2dii() +
  theme(
    legend.title = element_text()
  ) +
  facet_wrap(ald_sector ~ shock_year_arg, scales = "free") +
  labs(
    title = "Annual PD changes per sector and shock year",
    x = "Year",
    y = "PD change (% points)"
    ) 
```

## EL changes

```{r} 
data_losses <- data_el %>%
  filter(name %in% c("expected_loss_baseline", "expected_loss_shock")) %>%
  mutate(
    name = str_replace(name, "expected_loss_", "")
  )

ggplot(data_losses, aes(x = name, y = value, fill = value)) +
  geom_bar(stat = "identity", color = "grey") +
  scale_x_discrete(position = "top", labels = r2dii.plot::to_title) +
  scale_y_continuous(expand = expansion(mult = c(.1, 0)), labels = comma) +
  scale_fill_gradient2(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "green") %>% pull(.data$hex),
      midpoint = 0,
      labels = comma,
      name = "Expected loss\n(currency)"
    ) +
  facet_wrap(~ ald_sector, scales = "free") +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    strip.placement = "outside"
  ) +
  labs(
    title = "Expected loss per sector",
    subtitle = "For baseline and shock scenarios",
    y = "Expected Loss (currency)"
  )
```
```{r}
ggplot(data_losses, aes(x = name, y = el_as_perc_exposure, fill = el_as_perc_exposure)) +
  geom_bar(stat = "identity", color = "grey") +
  scale_x_discrete(position = "top", labels = r2dii.plot::to_title) +
  scale_y_continuous(expand = expansion(mult = c(.1, 0)), labels = percent) +
  scale_fill_gradient2(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "green") %>% pull(.data$hex),
      midpoint = 0,
      labels = percent,
      name = "Expected loss\n(% exposure)"
    ) +
  facet_wrap(~ ald_sector, scales = "free") +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    strip.placement = "outside"
  ) +
  labs(
    title = "Expected Loss as percentage of exposure per sector",
    subtitle = "For baseline and shock scenarios",
    y = "Expected Loss (% exposure)"
  )
```
```{r}
data_change <- data_el %>%
  filter(name %in% c("value_change")) 

p_abs <- ggplot(data_change, aes(x = name, y = value, fill = value)) +
  geom_bar(stat = "identity", color = "grey") +
  scale_x_discrete(position = "top", labels = r2dii.plot::to_title) +
  scale_y_continuous(expand = expansion(mult = c(.1, 0)), labels = comma) +
  scale_fill_gradient2(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "green") %>% pull(.data$hex),
      midpoint = 0,
      labels = comma,
      name = "Expected loss\n(currency)"
    ) +
  facet_wrap(~ ald_sector) +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    strip.placement = "outside"
  ) +
  labs(
    title = "Expected loss shock per sector",
    y = "Expected Loss\n(currency)"
  )

p_perc <- ggplot(data_change, aes(x = name, y = el_as_perc_exposure, fill = el_as_perc_exposure)) +
  geom_bar(stat = "identity", color = "grey") +
  scale_x_discrete(position = "top", labels = r2dii.plot::to_title) +
  scale_y_continuous(expand = expansion(mult = c(.1, 0)), labels = percent) +
  scale_fill_gradient2(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "green") %>% pull(.data$hex),
      midpoint = 0,
      labels = percent,
      name = "Expected loss\n(% exposure)"
    ) +
  facet_wrap(~ ald_sector) +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    strip.placement = "outside"
  ) +
  labs(
    title = "Expected loss shock as percentage of exposure per sector",
    y = "Expected Loss\n(% exposure)"
  )

p_abs / p_perc
```

## Equity Value changes

Total percentage

```{r}

data_plt <- data_val_change %>%
  group_by(portfolio_name) %>%
  select(portfolio_name, analysed_portfolio_percent_value_change) %>%
  distinct()

p_perc <- ggplot(data_plt, aes(x = portfolio_name, y = analysed_portfolio_percent_value_change)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 0) +
  theme_2dii() +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  ) +
  labs(
    title = "Analysed portfolio percentage value change",
    y = "Value change (% points)"
  )

data_plt <- data_val_change %>%
  group_by(portfolio_name) %>%
  select(portfolio_name, analysed_portfolio_absolute_value_change) %>%
  distinct()

p_abs <- ggplot(data_plt, aes(x = portfolio_name, y = analysed_portfolio_absolute_value_change)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_y_continuous(labels = scales::comma) +
  theme_2dii() +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  ) +
  labs(
    title = "Analysed portfolio absolute value change",
    y = "Value change (currency)"
  )

(p_perc + plot_spacer())/ (p_abs + plot_spacer())

```

Per sector

```{r}

data_plt <- data_val_change %>%
  group_by(ald_sector) %>%
  select(portfolio_name, sector_percent_value_change, sector_absolute_value_change, ald_sector) %>%
  distinct()

p_perc <- ggplot(data_plt, aes(x = portfolio_name, y = sector_percent_value_change, fill = sector_percent_value_change)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_gradient2(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "green") %>% pull(.data$hex),
      midpoint = 0,
      name = "Equity value change\n(% points)"
    ) +
  facet_wrap(ald_sector ~ .) +
  theme_2dii() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.line.x = element_blank()
  ) +
  labs(
    title = "Percentage value change per sector",
    y = "Value change\n(% points)"
  )

p_abs <- ggplot(data_plt, aes(x = portfolio_name, y = sector_absolute_value_change, fill = sector_absolute_value_change)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_gradient2(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "green") %>% pull(.data$hex),
      midpoint = 0,
      labels = scales::comma,
      name = "Equity value change\n(currency)"
    ) +
  facet_wrap(ald_sector ~ .) +
  theme_2dii() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.line.x = element_blank(),
    legend.text = element_text()
  ) +
  labs(
    title = "Absolute value change per sector",
    y = "Value change\n(currency)"
  )

p_perc / p_abs
```
Per technology

```{r}
ggplot(data_val_change, aes(x = portfolio_name, y = technology_percent_value_change, fill = technology_percent_value_change)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_gradient2(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "green") %>% pull(.data$hex),
      midpoint = 0,
      name = "Equity value\nchange (% points)"
    ) +
  facet_wrap(ald_sector ~ technology) +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.title.x = element_blank()
  ) +
  labs(
    title = "Percentage value change per technology",
    y = "Value change\n(% points)"
  )
```

```{r}
ggplot(data_val_change, aes(x = portfolio_name, y = technology_absolute_value_change, fill = technology_absolute_value_change)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_gradient2(
      low = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "red") %>% pull(.data$hex),
      high = r2dii.colours::palette_1in1000_plot %>% filter(.data$label == "green") %>% pull(.data$hex),
      midpoint = 0,
      labels = scales::comma,
      name = "Equity value\nchange (currency)"
    ) +
  facet_wrap(ald_sector ~ technology) +
  theme_2dii() +
  theme(
    legend.title = element_text(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.title.x = element_blank()
  ) +
  labs(
    title = "Absolute value change per technology",
    y = "Value change\n(currency)"
  )
```