---
title: "st_plots_using_functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{st_plots_using_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stress.test.plot.report)

granularity <-
  c(
    "ald_sector",
    "ald_business_unit"
  )

maturity_month_term_bridge_fp <-
  here::here("data", "maturity_month_term_bridge.csv")
portfolio_data_path <-
  here::here("data-raw/synthetic_portfolio.csv")
trisk_output_dir <- here::here("data-raw/synthetic_crispy")


trisk_start_year <- 2022
```

Read in the data. Put the outputs of a stress test run into 'data' folder of this project to run the code successfully.

```{r}
analysis_data <-
  load_input_plots_data(
    trisk_output_dir = trisk_output_dir,
    portfolio_data_path = portfolio_data_path,
    granularity = granularity,
    maturity_month_term_bridge_fp = maturity_month_term_bridge_fp,
    trisk_start_year = trisk_start_year
  )
```


Filter the trisk data to focus on a single geography/run

```{r}
focus_run_id <- "be4ce08a-6bc5-4724-a259-d5e2bd8fe01f"
focus_scenario_geography <- "Global"
active_tab <- "portfolio.portfolio_id"

analysis_data_single_run <- analysis_data |>
  dplyr::filter(
    .data$crispy.run_id == focus_run_id,
    .data$crispy.scenario_geography == focus_scenario_geography
  )
```

prepare data for plots

```{r}
data_exposure_plot <- prepare_for_exposure_plot(
  analysis_data_single_run,
  group_variable_char = active_tab,
  value_to_plot_char = "exposure_at_default"
)
data_performers_plot <- prepare_for_performers_plot(
  analysis_data_single_run,
  group_variable_vec = c(active_tab),
  performer_variable_char = "portfolio.ald_business_unit",
  performance_metric = "exposure_at_default",
  n_performers = 10
)
data_summary_plot <-
  prepare_for_summary_plot(
    analysis_data_single_run,
    group_variable_char = active_tab,
    value_to_plot_char = "exposure_at_default"
  )
data_el_plot <-
  prepare_for_el_plots(
    analysis_data_single_run,
    group_variable_char = active_tab,
    value_to_plot_char = "exposure_at_default"
  )
```


```{r}
data_pd <-
  prepare_for_pd_plot(
    analysis_data_single_run,
    group_variable_charvec = c("portfolio.ald_sector", "portfolio.term"),
    value_to_plot_char = "pd_difference"
  )

data_val_change <-
  prepare_for_value_change_plot(
    analysis_data_single_run,
    group_variable_charvec = c("portfolio.ald_sector", "portfolio.term"),
    value_to_plot_char = "shock"
  )
```

# Exposure plots

## Summary exposures per sector

Pie chart of exposures summary

```{r}
qplot_pie_summaries(data_summary_plot, agg_sum_name_chr = "group_sum")
```

## Risk-driving companies (based on exposure)

10 worst performers - coloured by sector

```{r fig2, fig.width = 8, fig.height=8}
qplot_worst_performers(data_performers_plot,
  x_var = "portfolio.ald_business_unit",
  y_var = "exposure_at_default",
  fill_var = active_tab,
  facet_var = active_tab
)
```

## Distributions

### Histograms and densities per sector

```{r}
qplot_histogram_density(data_exposure_plot)
```


### Boxplots and violin plots

Boxplots

```{r}
qplot_boxplot(data_exposure_plot)
```

Points distribution with violin

```{r}
qplot_violin(data_exposure_plot)
```

## PD changes

### Overall

Faceted - joint y-axis (easier to compare between sectors)

```{r}
qplot_pd_sector(data_pd)
```

```{r}
qplot_pd_sector(data_pd, common_y_axis = TRUE)
```


## EL changes

```{r fig3, fig.width = 8, fig.height = 8} 
qplot_el_sector(data_el_plot, use_exp_loss_types = c("baseline", "shock", "st_diff"))
```
```{r}
qplot_el_change_sector(data_el_plot)
```

## Equity Value changes

Total

```{r}
qplot_val_change(data_val_change)
```

Per sector

```{r} 
qplot_val_change_facet(data_val_change, facet_chr = "portfolio.ald_sector")
```

Per technology

```{r fig4, fig.width = 8, fig.height = 12}
# qplot_val_change_facet(data_val_change, facet_chr="ald_business_unit")
```
