---
title: "st_plots_using_functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{st_plots_using_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
devtools::load_all()
# library(stress.test.plot.report)


max_portfolio_granularity <-
  c("portfolio_id",
    "asset_type",
    "ald_sector",
    "ald_business_unit",
    "term")
max_crispy_granularity <-  c("run_id",
                             "scenario_geography",
                             "ald_sector",
                             "ald_business_unit",
                             "term")
portfolio_crispy_merge_cols <-
  c("ald_sector", "ald_business_unit", "term")

focus_run_id <- "NGFS2021_MESSAGE_B2DS_Global_no_carbon_tax"
focus_scenario_geography <- "Global"

maturity_month_term_bridge_fp <-
  here::here("data", "maturity_month_term_bridge.csv")
portfolio_data_path <-
  here::here("data-raw/synthetic_portfolio.csv")
crispy_outputs_dir <- here::here("data-raw/synthetic_crispy")


trisk_start_year <- 2022

active_tab <- "portfolio_id"
```

Read in the data. Put the outputs of a stress test run into 'data' folder of this project to run the code successfully.

```{r}
# load data
raw_portfolio_data <- load_portfolio_data(portfolio_data_path) 
maturity_month_term_bridge <- readr::read_csv(maturity_month_term_bridge_fp)

portfolio_data <-
  main_load_portfolio_data(
    raw_portfolio_data=raw_portfolio_data,
    maturity_month_term_bridge = maturity_month_term_bridge,
    max_portfolio_granularity = max_portfolio_granularity,
    trisk_start_year = trisk_start_year
  )

multi_crispy_data <- main_load_multi_crispy_data(crispy_outputs_dir = crispy_outputs_dir,
                                                 max_crispy_granularity = max_crispy_granularity)

trisk_run_params <- main_load_trisk_run_params(crispy_outputs_dir = crispy_outputs_dir)

analysis_data <-
  main_load_analysis_data(
    portfolio_data = portfolio_data,
    multi_crispy_data = multi_crispy_data,
    trisk_run_params = trisk_run_params,
    portfolio_crispy_merge_cols = portfolio_crispy_merge_cols
  )
```


Filter the trisk data to focus on a single geography/run

```{r}
analysis_data_single_run <- analysis_data |>
  dplyr::filter(.data$run_id == focus_run_id,
                .data$scenario_geography == focus_scenario_geography)
```

prepare data for plots

```{r}
data_exposure_plot <-
  prepare_for_exposure_plots(
    analysis_data_single_run,
    group_variable_vec = c(active_tab),
    agg_fn = sum
  )
data_performers_plot <- prepare_for_performers_plot(
  analysis_data_single_run,
  group_variable_vec = c(active_tab),
  performer_variable_char="ald_business_unit",
  performance_metric = "exposure_at_default",
  n_performers = 10
)
data_summary_plot <-
  prepare_for_summary_plot(
    analysis_data_single_run,
    group_variable_char = active_tab,
    value_to_plot_char = "exposure_at_default"
  )
data_el_plot <-
  prepare_for_el_plots(
    analysis_data_single_run,
    group_variable_char = active_tab,
    value_to_plot_char = "exposure_at_default"
  )
```


```{r}
data_pd <-
  prepare_for_pd_plot(
    analysis_data_single_run,
    group_variable_charvec = c("ald_sector", "term"),
    value_to_plot_char = "pd_difference_st"
  )
# data_pd_annual <-
#   prepare_for_pd_annual_plot(
#     analysis_data_single_run,
#     group_variable_char = c("ald_sector", "term", ""),
#     value_to_plot_char = "pd_difference_st"
#   )
data_val_change <-
  prepare_for_value_change_plot(
    analysis_data_single_run,
    group_variable_charvec = c("portfolio_id", "ald_sector"),
    value_to_plot_char = "shock"
  )
```


# Exposure plots

## Summary exposures per sector

Pie chart of exposures summary

```{r}
qplot_pie_summaries(data_summary_plot, agg_sum_name_chr="group_sum")
```

## Risk-driving companies (based on exposure)

10 worst performers - coloured by sector

```{r fig2, fig.width = 8, fig.height=8}
qplot_worst_performers(data_performers_plot, 
  x_var = "ald_business_unit",
  y_var = "exposure_at_default",
  fill_var = active_tab,
  facet_var = active_tab
  )

```

## Distributions

### Histograms and densities per sector

```{r}
qplot_histogram_density(data_exposure_plot) 
```


### Boxplots and violin plots

Boxplots

```{r}
qplot_boxplot(data_exposure_plot)
```

Points distribution with violin

```{r}
qplot_violin(data_exposure_plot)
```

## PD changes

### Overall

Faceted - joint y-axis (easier to compare between sectors)

```{r}
qplot_pd_sector(data_pd)
```

```{r}
qplot_pd_sector(data_pd, common_y_axis = TRUE)
```


### Annual
<!--
Faceted - joint y-axis (easier to compare between sectors)

```{r}
qplot_pd_sector(data_pd_annual, annual_pds = TRUE)
qplot_pd_sector(data_pd_annual, annual_pds = TRUE, common_y_axis = TRUE)
```
-->

## EL changes

```{r fig3, fig.width = 8, fig.height = 8} 
qplot_el_sector(data_el_plot, use_exp_loss_types= c("baseline", "shock", "st_diff"))
```
```{r}
qplot_el_change_sector(data_el_plot)
```

## Equity Value changes

Total

```{r}
qplot_val_change(data_val_change)
```

Per sector

```{r} 
qplot_val_change_facet(data_val_change, facet_chr="ald_sector")
```

Per technology

```{r fig4, fig.width = 8, fig.height = 12}
# qplot_val_change_facet(data_val_change, facet_chr="ald_business_unit")
```
