---
title: "st_plots_using_functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{st_plots_using_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
devtools::load_all()
library(stress.test.plot.report)
```

Read in the data. Put the outputs of a stress test run into 'data' folder of this project to run the code successfully.

```{r}
maturity_month_term_bridge_fp = here::here("data","maturity_month_term_bridge.csv")
maturity_month_term_bridge <- readr::read_csv(maturity_month_term_bridge_fp)

portfolio_data_path <- here::here("data-raw/synthetic_portfolio.csv")
portfolio_data <- load_portfolio_data(portfolio_data_path) |>
  map_portfolio_maturity_to_term(maturity_month_term_bridge, start_year=2022)

crispy_output_dir <- here::here("data-raw/synthetic_crispy")
multi_crispy <- load_multiple_crispy(crispy_output_dir)

analysis_data <- create_analysis_data(portfolio_data, multi_crispy)
```

# Exposure plots

## Summary exposures per sector

Pie chart of exposures summary

```{r}
qplot_pie_summaries(data_summary)
```

## Risk-driving companies (based on exposure)

10 worst performers - coloured by sector

```{r fig2, fig.width = 8, fig.height=8}
qplot_worst_performers(data_plot)
```

## Distributions

### Histograms and densities per sector

```{r}
qplot_histogram_density(data_plot) 
```


### Boxplots and violin plots

Boxplots

```{r}
qplot_boxplot(data_plot)
```

Points distribution with violin

```{r}
qplot_violin(data_plot)
```

## PD changes

### Overall

Faceted - joint y-axis (easier to compare between sectors)

```{r}
qplot_pd_sector(data_pd)
qplot_pd_sector(data_pd, common_y_axis = TRUE)
```

### Annual

Faceted - joint y-axis (easier to compare between sectors)

```{r}
qplot_pd_sector(data_pd_annual, annual_pds = TRUE)
qplot_pd_sector(data_pd_annual, annual_pds = TRUE, common_y_axis = TRUE)
```


## EL changes

```{r fig3, fig.width = 8, fig.height = 8} 
qplot_el_sector(data_el)
```
```{r}
qplot_el_change_sector(data_el)
```

## Equity Value changes

Total

```{r}
qplot_val_change(data_val_change)
```

Per sector

```{r} 
qplot_val_change_sector(data_val_change)
```

Per technology

```{r fig4, fig.width = 8, fig.height = 12}
qplot_val_change_technology(data_val_change)
```
